# **Practical Machine Learning Project**
## *Fernando Gonzalez Prada*
### *July 25th, 2015*


## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


## Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 


## Load the packages
```{r}
library(caret)
library(randomForest)
library(e1071)
```

## Read the data files into R data frames:

```{r}
train <- read.csv("pml-training.csv", row.names = 1, header = TRUE, na.strings = c("NA","","#DIV/0!"))
test <- read.csv("pml-testing.csv", row.names = 1, header = TRUE, na.strings = c("NA","","#DIV/0!"))
dim(train)
```
The training set consists of 19622 rows and 159 columns.

## Data Cleaning
The files are loaded mapping the NA, "" and ""#DIV/0!"" values into NA for missing values handling.
The are a lot of NA values present in multiple variables. To preserve the space of this report, the summary() command output is omitted, 

Remove the near zero variance predictors, to reduce the total size of the data frame
```{r}
nzv_cols <- nearZeroVar(train[, -159])
if(length(nzv_cols) > 0) 
    train <- train[, -nzv_cols]
dim(train)
```
Now we have 123 predictors

The next step involves deleting the remaining columns that contains values with NA.
These columns has almost all the values in NA, so we decided to delete the columns. 

```{r}
train = train[ , colSums(is.na(train)) == 0]
dim(train)
```
Finally, the data frame has 58 variables.


## Random Forest Model
After trying different models, the best results were obtained with Random Forest.
The solution involves fiting the algorithm performing 10-fold cross validation to predict the target variable "classe".

```{r}
set.seed(1000)

rfModel <- train(classe ~ ., 
               method = "rf", 
               data = train, 
               importance = T, 
               trControl = trainControl(method = "cv", number = 10))
rfModel
```
The accuracy of this Random Forest Model is 99%, quite impressive.

## Accuracy Chart
```{r}
plot(rfModel, ylim = c(0.95, 1))
```

## Final model and prediction
```{r}
rfModel$finalModel
```
The out of sample error is 0.04%

## Variable Importance
```{r}
rfImp = varImp(rfModel, scale = TRUE)
rfImp
```
The final model consists of 500 trees, and tried with 40 variables at each split. 
The 5 most important predictors are:

* raw_timestamp_part_1 
* roll_belt
* num_window
* pitch_forearm
* magnet_dumbbell_z

```{r}
plot(rfImp)
```


## Prediction
```{r}
pred <- predict(rfModel, test)
pred
```



## Generate the prediction files to submit to the autograder
```{r}
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("./prediction/problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
    }
}
pml_write_files(pred)

```



